sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2023-08-31 doc/paper-jss/paper-jss.Rmd
Init comments on fish models

We begin with a simple species distribution model of encounter probability of
Pacific Cod (TODO spp) from a trawl survey conducted off the west
coast of Vancouver Island, British Columbia, Canada. We will fit a logistic
regression of encounter probability with and without spatial random fields to
illustrate the importance of accounting for spatial correlation.

@@

We can also see that the AIC is considerably lower:

@@

@@

We can use `update()` to refit our model with four additional families:
delta-gamma, delta-lognormal, delta-Poisson-link-gamma, and
delta-Poisson-link-lognormal.

@@

We find that the anisotropic AR(1) is favoured. On inspecting the model output
with `summary()`, we notice that the spatial random field SD has become very
small once AR(1) spatiotemporal fields were added. We therefore consider
a model where these are turned off in the second component of the delta model:

@@

```{r dog-grid, echo=TRUE}
head(wcvi_grid, n = 3)
grid <- replicate_df(wcvi_grid, "year", time_values = unique(dat$year))
grid$log_depth <- log(grid$depth)
head(grid, n = 3)
```

>>>>

We begin with a simple species distribution model of encounter probability of
Pacific Cod (TODO spp) from a trawl survey conducted off the west
coast of Vancouver Island, British Columbia, Canada. This survey consists of spatially random sampling in each year, resulting in a georeferenced dataset. For compatibility with other packages, spatial random fields in \pkg{sdmTMB} can be toggled on and off. We will fit a logistic
regression of encounter probability with and without spatial random fields to
illustrate the importance of accounting for spatial correlation. In addition to the spatial random field, we include an intercept and a quadratic effect of depth on the probability of occurrence.

@@

We can also see that the AIC of the model with spatial random fields is considerably lower:

@@

<!-- Need brief description of dogfish data and model here?? -->

As a second example, we will construct a model of catch rates of North Pacific spiny dogfish from the same survey. This example extends the spatial model described previously by also including (1) spatiotemporal fields allowing the distribution in each year to be unique (modelled as an autoregressive process, allowing hotspots to persist through time), (2) a time varying intercept to allow each year to have a different mean catch rate, and (3) a smooth effect of ln(depth), allowing catch rates to vary non-linearly with depth. As catch rates are both skewed and zero inflated, we specify the response family as a Tweedie distribution [@tweedie1984].

@@

We can use `update()` to refit our model quickly without specifying all of the arguments. This is highlighted by exploring models with four additional families, each using a delta- or hurdle distribution:
delta-gamma, delta-lognormal, delta-Poisson-link-gamma, and
delta-Poisson-link-lognormal.

@@

<!-- Need description of anisotropy and why it's important here or somewhere else -->
We find that the anisotropic AR(1) is favoured. On inspecting the model output
with `summary()`, we notice that the spatial random field SD has become very
small once AR(1) spatiotemporal fields were added. We therefore consider
a model where these are turned off in the second component of the delta model:

@@

<!-- Propose moving this to a short paragraph `Prediction` to highlight modularity -->
```{r dog-grid, echo=TRUE}
head(wcvi_grid, n = 3)
grid <- replicate_df(wcvi_grid, "year", time_values = unique(dat$year))
grid$log_depth <- log(grid$depth)
head(grid, n = 3)
```
